---
title: "acad_thesis_final"
output: html_document
date: "2023-04-18"
---
```{r}
library(dplyr)
library(tidyverse)
library(epitab)
library(rstatix)
library(ggplot2)
library(broom)
library(tableone)
library(oddsratio)
library(flextable)
library(gtsummary)
library(naniar)
```

## Importing and cleaning data

Raw CSV file used instead due to column misalignment in the first dataset.

```{r}
acad_data = read_csv(file = "./data/ACAD-202301-Columbia-Wei-20230331-raw.csv") %>%
  janitor::clean_names() 
```

The CSV file was downloaded from the ACAD dataset. The dataset `acad_data` has 311 observations and 665 variables. The relevant variables were then selected for further cleaning.
 
```{r}
acad_clean = acad_data %>%
  select(
    "acad_id", "age","degree", "gender", "country", "generation", "redcap_data_access_group", "arrival_year",
    
    "parta_a201", "parta_a202", "parta_a203", "parta_a204", "parta_a205", "parta_a206", "parta_a207", "parta_a208", "parta_a209", "parta_a210", "parta_a211", "parta_a212", "parta_a213", "parta_a214", "parta_a2score",
    
    "memory", "orient", "judgment", "commun", "homehobb", "perscare", "cdr_sum", "cdrglob",
    
    "bills", "taxes", "shopping", "games", "stove", "mealprep", "events", "payattn", "remdates", "travel", "fas", 
    
    "ani_total", "veg_total",
    
    "clock1", "clock2", "clock3", "clock4", "clock5", "clock6", "clock7", "clock8", "clock9", "clock10", "clock11", "clock12", "clock13", "clock14", "clock15", "clock_t",
    
    "word1_t1", "word2_t1", "word3_t1", "word4_t1", "word5_t1", "word6_t1", "word7_t1", "word8_t1", "word9_t1", "word10_t1", "comt_t1","word1_t2", "word2_t2", "word3_t2", "word4_t2", "word5_t2", "word6_t2", "word7_t2", "word8_t2", "word9_t2", "word10_t2", "comt_t2", "word1_t3", "word2_t3", "word3_t3", "word4_t3", "word5_t3", "word6_t3", "word7_t3", "word8_t3", "word9_t3", "word10_t3", "comt_t3","word1_d", "word2_d", "word3_d", "word4_d", "word5_d", "word6_d", "word7_d", "word8_d", "word9_d", "word10_d", "comt_d", "item1", "item2", "item3", "item4", "item5", "item6", "item7", "item8", "item9", "item10", "word1_dr", "word2_dr", "word3_dr", "word4_dr", "word5_dr", "word6_dr", "word7_dr", "word8_dr","word9_dr", "word10_dr", "comt_dr",
    
    "casi_score", "casi2_score", "c14_overall", "c14_secg",
    
    "condition1", "condition2", "condition3", "condition4", "condition5", "condition6", "condition7", "condition8", "condition9", "condition10", "condition11", "condition12", "condition13", "condition14", "condition15"
  ) 
```

### Demographic Dataset

A dataset containing only demographic and covariate variables was created

```{r}
acad_dem = acad_clean %>%
  select("acad_id", "redcap_data_access_group", "age","degree", "gender", "country", "generation",  "arrival_year", "condition1", "condition2", "condition3", "condition4", "condition5", "condition6", "condition7", "condition8", "condition9", "condition10", "condition11", "condition12", "condition13", "condition14", "condition15")
```

Calculating immigration years; condition variables renamed 

```{r}
acad_dem = acad_dem %>%
  mutate(
  immigration_years = 2023 - arrival_year,
  
  hypertension = condition1,
  diabetes = condition2,
  cholesterol = condition3,
  myo_infarc = condition4,
  cong_heart_f = condition5,
  stroke = condition6, 
  head_injury = condition7,
  seizure = condition8,
  hiv = condition9,
  sleep_apnea = condition10,
  sleep_disorder = condition11,
  other_cog = condition12,
  covid = condition13,
  vac_covid = condition14,
  kidney = condition15,
  ) %>%
  
  select(-arrival_year, -condition1, -condition2, -condition3, -condition4, -condition5, -condition6, -condition7, -condition8, -condition9, -condition10, -condition11, -condition12, -condition13, -condition14, -condition15) 
```

Picking comorbidities: 

```{r}
acad_dem%>%
  group_by(hypertension) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(diabetes) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(cholesterol) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(myo_infarc) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(cong_heart_f) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(stroke) %>%
  summarise(n_obs = n())

acad_dem%>%
  group_by(head_injury) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(seizure) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(hiv) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(sleep_apnea) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(sleep_disorder) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(other_cog) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(covid) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(vac_covid) %>%
  summarise(n_obs = n())

acad_dem %>%
  group_by(kidney) %>%
  summarise(n_obs = n())
```

Selected the cardiovascular comorbidities (cholesterol, hypertension, diabetes, myo_infarc, cong_heart_f, stroke). Created cardiovascular risk score from comorbidities.  

```{r}
acad_dem = acad_dem %>%
  select(-head_injury, -seizure, -hiv, -sleep_apnea, -sleep_disorder, -other_cog, -covid, -vac_covid, -kidney) 

  acad_dem = acad_dem %>%
    replace_with_na(replace = list (
      cholesterol = c("55"),
      hypertension = c("55"),
      diabetes = c("55"),
      myo_infarc = c("55"),
      cong_heart_f = c("55"),
      stroke = c("55")
    )) %>%
  mutate(
    cardiovascular_risk = cholesterol + hypertension + diabetes + myo_infarc + cong_heart_f + stroke
  ) %>%
  mutate(
    cholesterol = factor(cholesterol, labels = c("No", "Yes")),
    hypertension = factor(hypertension, labels = c("No", "Yes")),
    diabetes = factor(diabetes, labels = c("No", "Yes")),
    myo_infarc = factor(myo_infarc, labels = c("No", "Yes")),
    cong_heart_f = factor(cong_heart_f, labels = c("No", "Yes")),
    stroke = factor(stroke, labels = c("No", "Yes")))

```

Remove participants with no information other than ACAD ID or recruitment site (n = 22)

```{r}
acad_dem <- acad_dem[!rowSums(is.na(acad_dem[, -c(1,2)])) == (ncol(acad_dem) -2),]
```

### Dietary Dataset

```{r}
acad_diet = acad_clean %>%
  select("acad_id","parta_a201", "parta_a202", "parta_a203", "parta_a204", "parta_a205", "parta_a206", "parta_a207", "parta_a208", "parta_a209", "parta_a210", "parta_a211", "parta_a212", "parta_a213", "parta_a214", "parta_a2score") %>%
  mutate(
  medi_score = as.factor(ifelse(parta_a2score < 5, "low",
                           ifelse(parta_a2score < 10, "middle",
                         "high"))),
  medi_score = factor(medi_score, levels = c("low", "middle", "high"))) %>%
  select("acad_id", "parta_a2score", "medi_score", everything())
```

Remove participants with no dietary information 

```{r}
acad_diet = acad_diet %>%
  drop_na(parta_a2score)
```
*Removed participants with parta_a2score = 0 since those participants had NA for all parta_a2 questions

(n total = 27)

```{r}
acad_diet = acad_diet %>%
  filter(, parta_a2score != 0 )
```


#### Distribution of parta_a2score 

Histogram of parta_a2score - since nicely distributed, no need to log transform
```{r}
hist_acad_diet_1 = hist(acad_diet$parta_a2score)
x_values = seq(min(acad_diet$parta_a2score), max(acad_diet$parta_a2score), length = 100)
y_values = dnorm(x_values, mean = mean(acad_diet$parta_a2score), sd = sd(acad_diet$parta_a2score))
y_values = y_values * diff(hist_acad_diet_1$mids[1:2]) * length(acad_diet$parta_a2score)

lines(x_values, y_values, lwd = 2)
```

Histogram of Medi Score
```{r}
acad_diet %>%
  ggplot(aes(x = medi_score)) +
  geom_bar()
```

### Cognitive Dataset

```{r}
acad_cog = acad_clean %>%
  select(
    "acad_id", "cdr_sum", "cdrglob", "fas", "ani_total", "veg_total","clock_t", "comt_t1","comt_t2", "comt_t3","comt_d", "comt_dr", "casi_score", "casi2_score", "c14_overall", "c14_secg"
  ) %>%
  
  mutate(
  cft = ani_total + veg_total,
  
  casi_score = ifelse(is.na(casi_score), 0, casi_score),
  casi2_score = ifelse(is.na(casi2_score), 0, casi2_score),
  casi_total = casi_score + casi2_score,
  casi_total = replace(casi_total, casi_total == 0, NA),
  
  comt_total = comt_t1 + comt_t2 + comt_t3 + comt_d,
  comt_total = ifelse(comt_total == 0, NA, comt_total
  ))%>%
    
  select(
"acad_id", "cdr_sum", "cdrglob", "fas","cft","clock_t", "comt_total", "comt_dr", "casi_total", "c14_overall", "c14_secg"
  )
```

Removing participants with zero entries for cognitive measures (only participants with at least one entry for a cognitive measure is included) (n = 22)

```{r}
acad_cog = acad_cog %>%
  mutate(
  cog_sum = rowSums(acad_cog[,c("cdr_sum", "cdrglob", "fas", "cft", "clock_t", "comt_total", "comt_dr", "casi_total", "c14_overall", "c14_secg")], na.rm = TRUE)
  )

acad_cog = acad_cog %>%
  filter(, cog_sum != 0) %>%
  select(-cog_sum)
```

#### CASI distribution 
```{r}
hist_acad_cog_1 = hist(acad_cog$casi_total)
```

Heavy left skew distribution; must log transform to create tertiles 

```{r}
acad_cog = acad_cog %>%
  mutate(
    casi_tertiles = cut(log(casi_total),
                        breaks = quantile(log(casi_total),
                                          probs = seq(0, 1, 1/3),
                                          na.rm = TRUE),
                        labels = c("1", "2", "3"),
                        include.lowest = TRUE,
                        include.highest = TRUE)
  )

acad_cog %>%
  ggplot(aes(x = casi_tertiles)) +
  geom_bar()
```

#### COMT distribution 
```{r}
hist_acad_cog_2 = hist(acad_cog$comt_total)
```

Heavy left skew distribution; must log transform to create tertiles 

```{r}
acad_cog = acad_cog %>%
  mutate(
    comt_tertiles = cut(log(comt_total),
                        breaks = quantile(log(comt_total),
                                          probs = seq(0, 1, 1/3),
                                          na.rm = TRUE),
                        labels = c("1", "2", "3"),
                        include.lowest = TRUE,
                        include.highest = TRUE)
  )

acad_cog %>%
  ggplot(aes(x = comt_tertiles)) +
  geom_bar()
```


### Merging all datasets

```{r}
dem_diet = inner_join(acad_dem, acad_diet, by = "acad_id")
thesis_final = inner_join(dem_diet, acad_cog, by = "acad_id")
```

Adding labels to final dataset and relabelling c14_overall

```{r}
thesis_final = thesis_final %>%
  mutate(
    degree = factor(degree, labels = c("Grade school or less", 
                                       "Some high school", 
                                       "High school or G.E.D.", 
                                       "Some college or technical school",
                                       "College graduate (Bachelor's degree)",
                                       "Graduate degree (Master's, Doctoral, or Professional degree)"))


,
    gender = factor(gender, labels = c("Male",
                                       "Female")),
    country = as.character(country),
    country = recode(country, 
                       "156" = "China", 
                       "410" = "Other", 
                       "124" = "Other", 
                       "344" = "Hong Kong", 
                       "158" = "Other", 
                       "360" = "Other", 
                       "458" = "Other", 
                       "888" = "Other", 
                       "704" = "Other" ),

    casi_tertiles = factor(casi_tertiles, labels = c(
      "Low", "Medium", "High"
      )),
    casi_tertiles = factor(casi_tertiles, levels = c(
      "Low", "Medium", "High"
      )),
    comt_tertiles = factor(comt_tertiles, labels = c(
      "Low", "Medium", "High"
      )),
    comt_tertiles = factor(comt_tertiles, levels = c(
      "Low", "Medium", "High"
      )),
    c14_relabels = case_when(
      c14_overall == 1 ~ 0,
      c14_overall %in% c(5, 6) ~ 1,
      c14_overall %in% c(2, 3, 4) ~ 2
    ),
    c14_secg = factor(c14_secg, labels = c(
      "No Dementia (includes MCI and SCC)", 
      "Dementia (NACC)"))
  ) %>%
  select(-c14_overall) %>%
  mutate(
    c14_overall = factor(c14_relabels, labels = c(
      "Normal",
      "MCI",
      "Dementia"
    ))
  ) %>%
  select(-c14_relabels)

skimr::skim(thesis_final)


write.csv(thesis_final, "thesis_final.csv", row.names = FALSE)
```

Final cleaned dataset is comprised of 303 observations and 43 variables - 3 character, 13 factor, and 27 numeric.

## Data Analysis

Exposure variables: 
* medi_score

Outcome variables:
* casi_total
* comt_total 

Secondary outcome variables
* c14_overall
* c14_secg
* cft 
* fas 


Covariates:
* age
* recruitment site (redcap_data_access_group)
* gender
* immigration years 
* education level
* cholesterol
* hypertension
* diabetes
* cardiovascular_risk

### Summary Tables 

#### Overall demographic and clinical characteristic table

```{r}
listVars1 = c("age", "redcap_data_access_group", "gender", "country", "immigration_years", "degree", "parta_a2score", "medi_score", "casi_total", "comt_total", "comt_dr", "fas","cft", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes", "cardiovascular_risk")

catVars = c("redcap_data_access_group", "gender", "country", "degree", "medi_score", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes")

table1 = CreateTableOne(vars = listVars1, data = thesis_final, factorVars = catVars)

print(table1, formatOption = list(big.mark = ","))
```

#### Demographic and clinical characteristics by Mediterranean diet score tertiles 

* Did not include casi and comt tertiles 

```{r}
listVars_2 = c("age", "redcap_data_access_group", "gender", "country", "immigration_years", "degree", "casi_total", "comt_total","comt_dr", "fas", "cft","c14_overall", "c14_secg",  "cholesterol", "hypertension", "diabetes", "cardiovascular_risk")

catVars = c("redcap_data_access_group", "gender", "country", "degree", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes")

table2 = CreateTableOne(vars = listVars_2, data = thesis_final, factorVars = catVars, strata = c("medi_score"))

print(table2, formatOptions = list(big.mark = ","))
```

#### Demographic, clinical, and dietary characteristics for subjects by casi tertiles 

```{r}
listVars_3 = c("age", "redcap_data_access_group", "gender", "country", "immigration_years", "degree", "parta_a2score","medi_score", "comt_total", "comt_dr", "fas", "cft","c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes", "cardiovascular_risk")

catVars = c("redcap_data_access_group", "gender", "country", "degree", "medi_score", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes")

table3 = CreateTableOne(vars = listVars_3, data = thesis_final, factorVars = catVars, strata = c("casi_tertiles"))

print(table3, formatOptions = list(big.mark = ","))
```

#### Demographic, clinical, and dietary characteristics for subjects by comt tertiles
```{r}
listVars_4 = c("age", "redcap_data_access_group", "gender", "country", "immigration_years", "degree", "parta_a2score", "medi_score", "casi_total", "comt_dr", "fas", "cft", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes", "cardiovascular_risk")

catVars = c("redcap_data_access_group", "gender", "country", "degree", "medi_score", "c14_overall", "c14_secg", "cholesterol", "hypertension", "diabetes")

table4 = CreateTableOne(vars = listVars_4, data = thesis_final, factorVars = catVars, strata = c("comt_tertiles"))

print(table4, formatOptions = list(big.mark = ","))
```

### Graphs 

#### Casi score by age 

```{r}
thesis_final %>%
  ggplot(aes(x = age, y = casi_total)) +
  geom_point() +
  stat_smooth(method = "loess", se = FALSE) +
  labs (
    x = "Age",
    y = "CASI Score"
  )
```

#### Comt score by age

```{r}
thesis_final %>%
  ggplot(aes(x = age, y = comt_total)) +
  geom_point() +
  stat_smooth(method = "loess", se = FALSE) +
  labs (
    x = "Age",
    y = "COMT Score"
  )
```

### Cognitive score by MeDi tertiles

```{r}
thesis_final %>%
  ggplot(aes(x = medi_score, y = casi_total)) +
  geom_boxplot()
```

### Cross sectional analysis 

Covariates: 
* age
* recruitment site (redcap_data_access_group)
* gender
* immigration years 
* education level
* cholesterol
* hypertension
* diabetes


#### Linear regression 

Linear regression between casi_total and medi_score
```{r}
#crude
linear1 = lm(casi_total ~ medi_score, data = thesis_final)
summary(linear1)
tbl_regression(linear1)

#model 1: age, gender, degree
linear1_model_1 = lm(casi_total ~ medi_score + age + gender + degree, data = thesis_final)
summary(linear1_model_1)
tbl_regression(linear1_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear1_model_2 = lm(casi_total ~ medi_score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = thesis_final)
summary(linear1_model_2)
tbl_regression(linear1_model_2)
```

Linear regression between casi_total and parta_a2score
```{r}
#crude
linear2 = lm(casi_total ~ parta_a2score, data = thesis_final)
summary(linear2)
tbl_regression(linear2)

#model 1: age, gender, degree
linear2_model_1 = lm(casi_total ~ parta_a2score + age + gender + degree, data = thesis_final)
summary(linear2_model_1)
tbl_regression(linear2_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear2_model_2 = lm(casi_total ~ parta_a2score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = thesis_final)
summary(linear2_model_2)
tbl_regression(linear2_model_2)
```

Linear regression between comt_total and medi_score
```{r}
#crude
linear3 = lm(comt_total ~ medi_score, data = thesis_final)
summary(linear3)
tbl_regression(linear3)

#model 1: age, gender, degree
linear3_model_1 = lm(comt_total ~ medi_score + age + gender + degree, data = thesis_final)
summary(linear3_model_1)
tbl_regression(linear3_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear3_model_2 = lm(comt_total ~ medi_score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = thesis_final)
summary(linear3_model_2)
tbl_regression(linear3_model_2)
```

Linear regression between comt_total and parta_a2score
```{r}
#crude
linear4 = lm(comt_total ~ parta_a2score, data = thesis_final)
summary(linear4)
tbl_regression(linear4)

#model 1: age, gender, degree
linear4_model_1 = lm(comt_total ~ parta_a2score + age + gender + degree, data = thesis_final)
summary(linear4_model_1)
tbl_regression(linear4_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear4_model_2 = lm(comt_total ~ parta_a2score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = thesis_final)
summary(linear4_model_2)
tbl_regression(linear4_model_2)
```

### Exploratory analysis

Logistic regression between clinical diagnosis (c14_secg) and medi_score

```{r}
#crude
thesis_final = thesis_final %>%
  mutate(
    c14_secg = relevel(c14_secg, "Dementia (NACC)")
  )

logit1 = glm(c14_secg ~ medi_score, data = thesis_final, family = "binomial")
summary(logit1)
tbl_regression(logit1, exponentiate = TRUE)
```


Linear regression between other cognitive tests and medi_score

FAS:
```{r}
#crude
linear5 = lm(fas ~ medi_score, data = thesis_final)
summary(linear5)
tbl_regression(linear5)
```

CFT:
```{r}
#crude
linear6 = lm(cft ~ medi_score, data = thesis_final)
summary(linear6)
tbl_regression(linear6)
```

#### Exclude dementia subjects from analysis and repeat association between medi_score and casi_toatl/comt_total 
```{r}
no_dementia = thesis_final 

no_dementia = no_dementia %>%
  filter(, c14_overall != "Dementia")
```

Linear regression between casi_total and medi_score
```{r}
#crude
linear7 = lm(casi_total ~ medi_score, data = no_dementia)
summary(linear7)
tbl_regression(linear7)

#model 1: age, gender, degree
linear7_model_1 = lm(casi_total ~ medi_score + age + gender + degree, data = no_dementia)
summary(linear7_model_1)
tbl_regression(linear7_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear7_model_2 = lm(casi_total ~ medi_score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = no_dementia)
summary(linear7_model_2)
tbl_regression(linear7_model_2)
```

Linear regression between casi_total and parta_a2score
```{r}
#crude
linear8 = lm(casi_total ~ parta_a2score, data = no_dementia)
summary(linear8)
tbl_regression(linear8)

#model 1: age, gender, degree
linear8_model_1 = lm(casi_total ~ parta_a2score + age + gender + degree, data = no_dementia)
summary(linear8_model_1)
tbl_regression(linear8_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear8_model_2 = lm(casi_total ~ parta_a2score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = no_dementia)
summary(linear8_model_2)
tbl_regression(linear8_model_2)
```

Linear regression between comt_total and medi_score
```{r}
#crude
linear9 = lm(comt_total ~ medi_score, data = no_dementia)
summary(linear9)
tbl_regression(linear9)

#model 1: age, gender, degree
linear9_model_1 = lm(comt_total ~ medi_score + age + gender + degree, data = no_dementia)
summary(linear9_model_1)
tbl_regression(linear9_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear9_model_2 = lm(comt_total ~ medi_score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = no_dementia)
summary(linear9_model_2)
tbl_regression(linear9_model_2)
```

Linear regression between comt_total and parta_a2score
```{r}
#crude
linear10= lm(comt_total ~ parta_a2score, data = no_dementia)
summary(linear10)
tbl_regression(linear10)

#model 1: age, gender, degree
linear10_model_1 = lm(comt_total ~ parta_a2score + age + gender + degree, data = no_dementia)
summary(linear10_model_1)
tbl_regression(linear10_model_1)

#model 2: age, gender, degree, redcap_data_access_group, immigration_years, cholesterol, diabetes, hypertension
linear10_model_2 = lm(comt_total ~ parta_a2score + age + redcap_data_access_group + gender + immigration_years + degree + cholesterol + hypertension + diabetes, data = no_dementia)
summary(linear10_model_2)
tbl_regression(linear10_model_2)
```

#### Individual food components and cognition association 

```{r}
#parta_a201: Olive oil as main cooking fat?
parta_a201_fit_casi = lm(casi_total ~ parta_a201, data = thesis_final)
summary(parta_a201_fit_casi)
tbl_regression(parta_a201_fit_casi)

parta_a201_fit_comt = lm(comt_total ~ parta_a201, data = thesis_final)
summary(parta_a201_fit_comt)
tbl_regression(parta_a201_fit_comt)

#parta_a202: More than 4 tablespoons of olive oil?
parta_a202_fit_casi = lm(casi_total ~ parta_a202, data = thesis_final)
summary(parta_a202_fit_casi)
tbl_regression(parta_a202_fit_casi)

parta_a202_fit_comt = lm(comt_total ~ parta_a202, data = thesis_final)
summary(parta_a202_fit_comt)
tbl_regression(parta_a202_fit_comt)

#parta_a203: 2 or more servings of vegetable in a day?
parta_a203_fit_casi = lm(casi_total ~ parta_a203, data = thesis_final)
summary(parta_a203_fit_casi)
tbl_regression(parta_a203_fit_casi)

parta_a203_fit_comt = lm(comt_total ~ parta_a203, data = thesis_final)
summary(parta_a203_fit_comt)
tbl_regression(parta_a203_fit_comt)

#parta_a204: 3 or more servings of fruit per day?
parta_a204_fit_casi = lm(casi_total ~ parta_a204, data = thesis_final)
summary(parta_a204_fit_casi)
tbl_regression(parta_a204_fit_casi)

parta_a204_fit_comt = lm(comt_total ~ parta_a204, data = thesis_final)
summary(parta_a204_fit_comt)
tbl_regression(parta_a204_fit_comt)

#parta_a205: Butter, margarine, or cream less than once daily?
parta_a205_fit_casi = lm(casi_total ~ parta_a205, data = thesis_final)
summary(parta_a205_fit_casi)
tbl_regression(parta_a205_fit_casi)

parta_a205_fit_comt = lm(comt_total ~ parta_a205, data = thesis_final)
summary(parta_a205_fit_comt)
tbl_regression(parta_a205_fit_comt)

#parta_a206: Beef, pork, or other meat less than daily?
parta_a206_fit_casi = lm(casi_total ~ parta_a206, data = thesis_final)
summary(parta_a206_fit_casi)
tbl_regression(parta_a206_fit_casi)

parta_a206_fit_comt = lm(comt_total ~ parta_a206, data = thesis_final)
summary(parta_a206_fit_comt)
tbl_regression(parta_a206_fit_comt)

#parta_a207: Sweetened/carbonated beverages less often than daily?
parta_a207_fit_casi = lm(casi_total ~ parta_a207, data = thesis_final)
summary(parta_a207_fit_casi)
tbl_regression(parta_a207_fit_casi)

parta_a207_fit_comt = lm(comt_total ~ parta_a207, data = thesis_final)
summary(parta_a207_fit_comt)
tbl_regression(parta_a207_fit_comt)

#parta_a208: At least 7 glasses of wine per week?
parta_a208_fit_casi = lm(casi_total ~ parta_a208, data = thesis_final)
summary(parta_a208_fit_casi)
tbl_regression(parta_a208_fit_casi)

parta_a208_fit_comt = lm(comt_total ~ parta_a208, data = thesis_final)
summary(parta_a208_fit_comt)
tbl_regression(parta_a208_fit_comt)

#parta_a209: 3 or more servings of lentils, peas, beans, soybeans, or tofu/week?
parta_a209_fit_casi = lm(casi_total ~ parta_a209, data = thesis_final)
summary(parta_a209_fit_casi)
tbl_regression(parta_a209_fit_casi)

parta_a209_fit_comt = lm(comt_total ~ parta_a209, data = thesis_final)
summary(parta_a209_fit_comt)
tbl_regression(parta_a209_fit_comt)

#parta_a210: 3 or more servings of fish/shellfish per week?
parta_a210_fit_casi = lm(casi_total ~ parta_a210, data = thesis_final)
summary(parta_a210_fit_casi)
tbl_regression(parta_a210_fit_casi)

parta_a210_fit_comt = lm(comt_total ~ parta_a210, data = thesis_final)
summary(parta_a210_fit_comt)
tbl_regression(parta_a210_fit_comt)

#parta_a211: Fewer than 2 servings per week of sweets/pastries?
parta_a211_fit_casi = lm(casi_total ~ parta_a211, data = thesis_final)
summary(parta_a211_fit_casi)
tbl_regression(parta_a211_fit_casi)

parta_a211_fit_comt = lm(comt_total ~ parta_a211, data = thesis_final)
summary(parta_a211_fit_comt)
tbl_regression(parta_a211_fit_comt)

#parta_a212: 3 or more servings of tree nuts per week?
parta_a212_fit_casi = lm(casi_total ~ parta_a212, data = thesis_final)
summary(parta_a212_fit_casi)
tbl_regression(parta_a212_fit_casi)

parta_a212_fit_comt = lm(comt_total ~ parta_a212, data = thesis_final)
summary(parta_a212_fit_comt)
tbl_regression(parta_a212_fit_comt)

#parta_a213: Generally eat more chicken/turkey/rabbit compared to veal/pork/hamburger/sausage?
parta_a213_fit_casi = lm(casi_total ~ parta_a213, data = thesis_final)
summary(parta_a213_fit_casi)
tbl_regression(parta_a213_fit_casi)

parta_a213_fit_comt = lm(comt_total ~ parta_a213, data = thesis_final)
summary(parta_a213_fit_comt)
tbl_regression(parta_a213_fit_comt)

#parta_a214: Season vegetables, pasta, rice iwth tomato, onion, leek, or garlic sauce 2 or more times per week?
parta_a214_fit_casi = lm(casi_total ~ parta_a214, data = thesis_final)
summary(parta_a214_fit_casi)
tbl_regression(parta_a214_fit_casi)

parta_a214_fit_comt = lm(comt_total ~ parta_a214, data = thesis_final)
summary(parta_a214_fit_comt)
tbl_regression(parta_a214_fit_comt)
```

#### Excluding dementia - Individual food components and cognitive assocation

```{r}
parta_a201_fit_casi = lm(casi_total ~ parta_a201, data = no_dementia)
summary(parta_a201_fit_casi)
tbl_regression(parta_a201_fit_casi)

parta_a201_fit_comt = lm(comt_total ~ parta_a201, data = no_dementia)
summary(parta_a201_fit_comt)
tbl_regression(parta_a201_fit_comt)

#parta_a202: More than 4 tablespoons of olive oil?
parta_a202_fit_casi = lm(casi_total ~ parta_a202, data = no_dementia)
summary(parta_a202_fit_casi)
tbl_regression(parta_a202_fit_casi)

parta_a202_fit_comt = lm(comt_total ~ parta_a202, data = no_dementia)
summary(parta_a202_fit_comt)
tbl_regression(parta_a202_fit_comt)

#parta_a203: 2 or more servings of vegetable in a day?
parta_a203_fit_casi = lm(casi_total ~ parta_a203, data = no_dementia)
summary(parta_a203_fit_casi)
tbl_regression(parta_a203_fit_casi)

parta_a203_fit_comt = lm(comt_total ~ parta_a203, data = no_dementia)
summary(parta_a203_fit_comt)
tbl_regression(parta_a203_fit_comt)

#parta_a204: 3 or more servings of fruit per day?
parta_a204_fit_casi = lm(casi_total ~ parta_a204, data = no_dementia)
summary(parta_a204_fit_casi)
tbl_regression(parta_a204_fit_casi)

parta_a204_fit_comt = lm(comt_total ~ parta_a204, data = no_dementia)
summary(parta_a204_fit_comt)
tbl_regression(parta_a204_fit_comt)

#parta_a205: Butter, margarine, or cream less than once daily?
parta_a205_fit_casi = lm(casi_total ~ parta_a205, data = no_dementia)
summary(parta_a205_fit_casi)
tbl_regression(parta_a205_fit_casi)

parta_a205_fit_comt = lm(comt_total ~ parta_a205, data = no_dementia)
summary(parta_a205_fit_comt)
tbl_regression(parta_a205_fit_comt)

#parta_a206: Beef, pork, or other meat less than daily?
parta_a206_fit_casi = lm(casi_total ~ parta_a206, data = no_dementia)
summary(parta_a206_fit_casi)
tbl_regression(parta_a206_fit_casi)

parta_a206_fit_comt = lm(comt_total ~ parta_a206, data = no_dementia)
summary(parta_a206_fit_comt)
tbl_regression(parta_a206_fit_comt)

#parta_a207: Sweetened/carbonated beverages less often than daily?
parta_a207_fit_casi = lm(casi_total ~ parta_a207, data = no_dementia)
summary(parta_a207_fit_casi)
tbl_regression(parta_a207_fit_casi)

parta_a207_fit_comt = lm(comt_total ~ parta_a207, data = no_dementia)
summary(parta_a207_fit_comt)
tbl_regression(parta_a207_fit_comt)

#parta_a208: At least 7 glasses of wine per week?
parta_a208_fit_casi = lm(casi_total ~ parta_a208, data = no_dementia)
summary(parta_a208_fit_casi)
tbl_regression(parta_a208_fit_casi)

parta_a208_fit_comt = lm(comt_total ~ parta_a208, data = no_dementia)
summary(parta_a208_fit_comt)
tbl_regression(parta_a208_fit_comt)

#parta_a209: 3 or more servings of lentils, peas, beans, soybeans, or tofu/week?
parta_a209_fit_casi = lm(casi_total ~ parta_a209, data = no_dementia)
summary(parta_a209_fit_casi)
tbl_regression(parta_a209_fit_casi)

parta_a209_fit_comt = lm(comt_total ~ parta_a209, data = no_dementia)
summary(parta_a209_fit_comt)
tbl_regression(parta_a209_fit_comt)

#parta_a210: 3 or more servings of fish/shellfish per week?
parta_a210_fit_casi = lm(casi_total ~ parta_a210, data = no_dementia)
summary(parta_a210_fit_casi)
tbl_regression(parta_a210_fit_casi)

parta_a210_fit_comt = lm(comt_total ~ parta_a210, data = no_dementia)
summary(parta_a210_fit_comt)
tbl_regression(parta_a210_fit_comt)

#parta_a211: Fewer than 2 servings per week of sweets/pastries?
parta_a211_fit_casi = lm(casi_total ~ parta_a211, data = no_dementia)
summary(parta_a211_fit_casi)
tbl_regression(parta_a211_fit_casi)

parta_a211_fit_comt = lm(comt_total ~ parta_a211, data = no_dementia)
summary(parta_a211_fit_comt)
tbl_regression(parta_a211_fit_comt)

#parta_a212: 3 or more servings of tree nuts per week?
parta_a212_fit_casi = lm(casi_total ~ parta_a212, data = no_dementia)
summary(parta_a212_fit_casi)
tbl_regression(parta_a212_fit_casi)

parta_a212_fit_comt = lm(comt_total ~ parta_a212, data = no_dementia)
summary(parta_a212_fit_comt)
tbl_regression(parta_a212_fit_comt)

#parta_a213: Generally eat more chicken/turkey/rabbit compared to veal/pork/hamburger/sausage?
parta_a213_fit_casi = lm(casi_total ~ parta_a213, data = no_dementia)
summary(parta_a213_fit_casi)
tbl_regression(parta_a213_fit_casi)

parta_a213_fit_comt = lm(comt_total ~ parta_a213, data = no_dementia)
summary(parta_a213_fit_comt)
tbl_regression(parta_a213_fit_comt)

#parta_a214: Season vegetables, pasta, rice iwth tomato, onion, leek, or garlic sauce 2 or more times per week?
parta_a214_fit_casi = lm(casi_total ~ parta_a214, data = no_dementia)
summary(parta_a214_fit_casi)
tbl_regression(parta_a214_fit_casi)

parta_a214_fit_comt = lm(comt_total ~ parta_a214, data = no_dementia)
summary(parta_a214_fit_comt)
tbl_regression(parta_a214_fit_comt)
```

#### Excluding dementia - Linear regression between other cognitive tests and medi_score

FAS:
```{r}
#crude
linear11 = lm(fas ~ medi_score, data = no_dementia)
summary(linear11)
tbl_regression(linear11)
```

CFT:
```{r}
#crude
linear12 = lm(cft ~ medi_score, data = no_dementia)
summary(linear12)
tbl_regression(linear12)
```

#### Excluding dementia - Logistic regression between casi_total/comt_total and Normal/MCI

```{r}
linear13 = lm(casi_total ~ c14_overall, data = no_dementia)
summary(linear13)
tbl_regression(linear13)

linear14 = lm(comt_total ~ c14_overall, data = no_dementia)
summary(linear14)
tbl_regression(linear14)
```

Table of avg CASI and COMT score stratified by Normal/MCI
```{r}
listVars_5 = c("casi_total", "comt_total")

catVars = c("c14_overall")

table5 = CreateTableOne(vars = listVars_5, data = no_dementia, factorVars = catVars, strata = c("c14_overall"))

print(table5, formatOptions = list(big.mark = ","))
```





